figure sandbox

### 2023-11-10 untargeted/paruntargeted_sandbox.Rmd
Adding contour plot fitting and kmeans/RF things

# polarity <- "pos"
# polarity <- "neg"
# amendment <- "Amm"
# amendment <- "Nit"
# amendment <- "GMP"
# amendment <- "Arg"
# IS_type <- "noIS"
# IS_type <- "wIS"

parametadata_all <- read_csv("metadata/parametadata.csv") %>% 
  mutate(startime=factor(startime, levels=c("Morn", "Eve"))) %>%
  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
  mutate(timepoint=factor(timepoint, levels=paste0("T", c(0,1,3,10,26,73))))
parametadata <- parametadata_all %>%
  filter(polarity==!!polarity) %>%
  filter(amendment==!!amendment) %>%
  filter(IS_type==!!IS_type | samp_type=="Std")
all_stans <- read_csv("metadata/all_stans.csv")

@@

```{r kmeans clustering of isotope signals}

```

>>>>

polarity <- "pos"
# polarity <- "neg"
amendment <- "Amm"
# amendment <- "Nit"
# amendment <- "GMP"
# amendment <- "Arg"
IS_type <- "noIS"
# IS_type <- "wIS"

parametadata_all <- read_csv("metadata/parametadata.csv") %>% 
  mutate(startime=factor(startime, levels=c("Morn", "Eve"))) %>%
  mutate(depth=factor(depth, levels=c("Surf", "Deep"))) %>%
  mutate(timepoint=factor(timepoint, levels=paste0("T", c(0,1,3,10,26,73))))
parametadata <- parametadata_all %>%
  filter(polarity==!!polarity) %>%
  filter(amendment==!!amendment) %>%
  filter(IS_type==!!IS_type | samp_type=="Std")
manual_typing <- readxl::read_excel("metadata/manual_typing.xlsx") %>%
  mutate(manual_type=ifelse(is.na(manual_type), "Other", manual_type)) %>%
  add_row(compound_name="Thymidine", manual_type="Nucleobases +")
all_stans <- read_csv("metadata/all_stans.csv") %>%
  left_join(manual_typing)

@@

```{r logistic curve fitting contour plot and curve fitting}
glm_coef_data <- all_concs %>%
  left_join(parametadata_all) %>%
  filter(amendment=="Amm") %>%
  filter(samp_type=="Smp") %>%
  filter(!str_detect(filename, "Blk")) %>%
  mutate(label_type=ifelse(labeled_n_count==0, "Unlabeled", "Labeled")) %>%
  select(compound_name, filename, label_type, nM_in_env, timepoint_num, depth, startime) %>%
  pivot_wider(names_from = label_type, values_from=nM_in_env, values_fn = sum) %>%
  mutate(frac_labeled=Labeled/(Labeled+Unlabeled)) %>%
  nest(data=-c(compound_name, depth)) %>%
  mutate(nls_test=map(data, ~broom::tidy(glm(.x$frac_labeled~.x$timepoint_num, family=quasibinomial)))) %>%
  unnest(nls_test) %>%
  select(compound_name, depth, data, term, estimate) %>%
  pivot_wider(names_from = term, values_from = estimate) %>%
  dplyr::rename(intercept=`(Intercept)`, slope=`.x$timepoint_num`) %>%
  mutate(intercept=1-(exp(-intercept)/(1+exp(-intercept)))) %>%
  left_join(manual_typing)
```

```{r kmeans clustering of isotope signals}
frac_labeled_df <- all_areas %>%
  left_join(parametadata_all) %>%
  filter(amendment=="Amm") %>%
  filter(samp_type=="Smp") %>%
  filter(!str_detect(filename, "Blk")) %>%
  filter(polarity=="pos") %>%
  mutate(label_type=ifelse(labeled_n_count==0, "Unlabeled", "Labeled")) %>%
  select(compound_name, shortname, label_type, area, timepoint_num, depth, startime) %>%
  pivot_wider(names_from = label_type, values_from=area, values_fn = sum) %>%
  mutate(frac_labeled=Labeled/(Labeled+Unlabeled)) %>%
  ungroup() %>%
  arrange(depth, startime, timepoint_num) %>%
  mutate(shortname=fct_inorder(shortname))

frac_labeled_df %>%
  filter(compound_name=="L-Glutamine") %>%
  ggplot() +
  geom_point(aes(x=timepoint_num, y=frac_labeled)) +
  facet_grid(depth~startime)

frac_mat <- frac_labeled_df %>%
  select(compound_name, shortname, frac_labeled) %>%
  pivot_wider(names_from=shortname, values_from=frac_labeled) %>%
  mutate(across(everything(), function(x)ifelse(is.na(x), 0, x))) %>%
  # filter(compound_name=="L-Glutamine") %>%
  # filter(is.na(`210916_Smp_Amm-Deep-Eve-T0-A-noIS-pos.mzML`)) %>%
  column_to_rownames("compound_name")

# heatmaply::heatmaply(frac_mat, Colv=FALSE)
kmeans_obj <- kmeans(frac_mat, centers = 4)
mapping <- `names<-`(1:6, order(kmeans_obj$size, decreasing = TRUE))
old_names <- names(kmeans_obj$cluster)
kmeans_obj$cluster <- mapping[as.character(kmeans_obj$cluster)]
names(kmeans_obj$cluster) <- old_names
kmeans_obj$size <- sort(kmeans_obj$size, decreasing = TRUE)
table(kmeans_obj$cluster)

data.frame(cluster=kmeans_obj$cluster) %>%
  # split(.$cluster)
  mutate(cluster=paste0("Cluster ", cluster, " n=", kmeans_obj$size[cluster])) %>%
  rownames_to_column("compound_name") %>%
  left_join(frac_labeled_df) %>%
  # ggplot() +
  # geom_boxplot(aes(x=timepoint_num, y=frac_labeled, color=startime, group=interaction(timepoint_num, startime))) +
  # facet_grid(depth~cluster)
  group_by(cluster, shortname) %>%
  summarise(mean_labeled=mean(frac_labeled, na.rm=TRUE)) %>%
  left_join(parametadata) %>%
  ggplot() +
  geom_point(aes(x=timepoint_num, y=mean_labeled, color=startime)) +
  facet_grid(depth~cluster)

library(randomForest)
library(rpart)
# library(rpart.plot)

(rfmodel <- randomForest(frac_mat, y=factor(kmeans_obj$cluster), ntree = 5000, 
                        importance=TRUE, proximity = TRUE, keep.forest = TRUE))
# varImpPlot(rfmodel, n.var = 60)
```
